The effects of binarity on planet occurrence rates measured by transit surveys

### 2017-12-19 paper/main.tex
paper: improve text for readability

\documentclass[12pt,modern]{aastex61}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{comment}
\usepackage{grffile} % checks for multiple dots in figure filenames
\usepackage{empheq} % for boxing equations
\newcommand*\widefbox[1]{\fbox{\hspace{1em}#1\hspace{1em}}}

@@

\newcommand{\pp}{\mathcal{P}}
\newcommand{\ps}{\mathcal{S}}
\renewcommand{\a}{_{\rm a}}
\newcommand{\s}{_{\rm s}}
\renewcommand{\d}{_{\rm d}}
\begin{document}

\title{ The effects of binarity on planet occurrence rates measured by transit 
surveys}
%
\correspondingauthor{L. Bouma}
\email{luke@astro.princeton.edu}
%
\author{L. G. Bouma}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author{J. N. Winn}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author{K. Masuda}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
%
%
\begin{abstract}
%
This study develops simplified models of signal-to-noise limited transit 
surveys, in order to clarify the biases that stellar binarity introduces
to occurrence rate measurements.
Particular attention is paid to dilution of observed planet radii, errors in 
star counts, and detection efficiencies.
Free parameters in our models include the true radius distribution of planets, 
and the relative numbers of planets orbiting secondaries, primaries, and 
single stars.
In the simplest possible case, all stellar systems are either single or twin 
binaries, and all planets are identical. If so, then we find that ignoring 
binarity leads to an underestimate of the occurrence at the true radius by a 
multiplicative factor of $0.76$ (assuming a 10\% twin binary fraction).
Using more realistic models for the stellar population and planetary 
radii, we find that the occurrence of Earth-sized planets can be overestimated 
by at most $\approx 50\%$ when ignoring binarity~--~at present far smaller 
than systematic uncertainties on real $\eta_\oplus$ measurements.
We quantify how gaps in the intrinsic planet radius distribution are filled 
in when binarity is neglected.
Our models suggest that for most of the observed radius distribution 
($r>2r_\oplus$), ignoring binarity only leads to errors in inferred occurrence 
rates at the $\lesssim 10\%$ level.
One corollary is that binarity is unlikely to strongly influence the HJ rate 
discrepancy.
%
\end{abstract}
%
\keywords{
    methods: data analysis ---
    planets and satellites: detection ---
    surveys}
%
%

@@

To compute an occurrence, at every planetary and stellar property the 
astronomers choose the stars (among those initially surveyed) around which the 
planets of interest appeared to be searchable.
Correcting for the geometric transit probability $p_{\rm tra}$, they report an 
apparent rate density,
\begin{equation}
\Gamma\a(\pp\a, \ps\a) = \frac{n_{\rm det}(\pp\a, \ps\a)}{N_{\rm s,a}(\pp\a, 
    \ps\a)} \times \frac{1}{p_{\rm tra}(\pp\a, \ps\a)}.
\label{eq:defn_apparent_rate_density}
\end{equation}
where $\pp\a$, $\ps\a$ are the apparent planetary and stellar parameters.
The quantity $N_{\rm s,a}(\pp\a, \ps\a)$ is the number of unresolved 
point-sources on the sky that appear to be searchable, and
$n_{\rm det}(\pp\a, \ps\a)$ is the number of detected planets, per unit 
$\pp\a$ and $\ps\a$.

@@

An immediate complication is that, due to dynamical stability or some 
aspect of planet formation, the occurrence rate of planets might differ 
between binary and single-star systems.
If ``occurrence rate'' is defined as the mean number of planets within 
set radius and period bounds per star in a given mass interval, it must 
implicitly marginalize over stellar multiplicity.
This means marginalizing over ``occurrence rates in single star systems'', 
``occurrence rates about primaries'', and
``occurrence rates about secondaries''
\citep[see][]{wang_occurrence_2015}.

Outside of astrophysical differences, every term in 
Eq.~\ref{eq:defn_apparent_rate_density} is observationally-biased.
There are errors in $n_{\rm det}(\pp\a,\ps\a)$ due to planet radius 
misclassification.
There are errors in $N_{{\rm s,a}}(\pp\a,\ps\a)$ because an apparently 
searchable point on the sky might include two searchable stars.
There are errors in $p_{\rm tra}(\pp\a,\ps\a)$ because stars in binaries may 
have different masses and radii than assumed in the single-star case.
Binarity might also affect the observers' ability to correctly select 
searchable stars (but see Sec.~\ref{sec:model_1} for discussion of this 
issue).

Correcting for binarity's observational biases is non-trivial.
For instance, in counting the number of selected stars, even after realizing 
that binaries count as two stars, one must note that the multiplicity fraction 
of {\it selected} stars is greater than that of a volume limited 
sample.
This is the familiar Malmquist bias: binaries are selected out to larger 
distances than single stars because they are more luminous.
As a separate challenge, finding the correct number of detected planets in a 
radius bin requires knowledge of the true planetary radii.
Observers deduce apparent radii.
In binaries, the apparent and true radii differ because of diluting flux, 
and possibly because the planet is assumed to orbit the wrong star
\citep[\textit{e.g.},][]{furlan_kepler_2017}.

To gain intuition for the many observational biases at play,
we consider a set of idealized transit surveys:
\begin{itemize}
    \item Model \#1: fixed stars, fixed planets, twin binaries;
    \item Model \#2: fixed primaries, power-law planets, varying secondaries;
    \item Model \#3: fixed primaries, broken power-law planets and varying 
    secondaries.
\end{itemize}
In Sec.~\ref{sec:definitions}, we clarify our terminology and state
assumptions that apply throughout this study.
We then present our transit survey models in 
Secs.~\ref{sec:model_1}-\ref{sec:model_3}, where 
each subsection corresponds to a model listed above.
We interpret these calculations throughout, and in 
Sec.~\ref{sec:discussion} connect them to topical questions in 
the interpretation of transit survey occurrence rates.
In particular, we mention the ``hot Jupiter rate discrepancy'', the relevance 
towards measurements of $\eta_\oplus$, and the implications for detailed 
study of \citet{fulton_california-_2017}'s recently discovered ``valley''.
We conclude in Sec.~\ref{sec:conclusion}.

@@

\section{Definitions and Preliminaries}
\label{sec:definitions}

\subsection{Occurrence rate, occurrence rate density}
Define the occurrence rate density, $\Gamma$, as the expected number of 
planets per star per unit of planetary and stellar phase space.
Since we will mainly be concerned with the rate density's dependence on 
planetary radii $r$, we write
\begin{equation}
\Gamma(r) = \frac{{\rm d}\Lambda}{{\rm d}r},
\label{eq:rate_density_defn}
\end{equation}
where $\Lambda$ is the occurrence rate.
In this notation, ``the occurrence rate of planets of a particular size'' 
translates to an integral of Eq.~\ref{eq:rate_density_defn}, evaluated over a 
radius interval.

The above definition implicitly marginalizes the rate density over stellar 
multiplicity.
Considering only single and binary systems, the rate density for any selected 
sample of stars is a weighted sum of rate densities for each system type:
\begin{equation}
N_{\rm tot} \Gamma_{\rm sel}(r) =
N_0 \Gamma_0(r) +
N_1 \Gamma_1(r) +
N_2 \Gamma_2(r),
\label{eq:rate_density_marginalized}
\end{equation}
where $i=0$ corresponds to single stars, $i=1$ to primaries of 
binaries, and $i=2$ to secondaries of binaries.
$N_{\rm tot} = \sum_i N_i$ is the total number of selected stars, and 
$N_0,N_1,N_2$ are the number of selected single stars, primaries, and 
secondaries. For our purposes binary systems are not identifiable, and so 
$N_1=N_2$.

The rate density for each type of star, $\Gamma_i(r)$, can be written as 
the product of a shape function and a constant:
\begin{equation}
\Gamma_i(r) = \frac{{\rm d}\Lambda_i}{{\rm d}r} = Z_i f_i(r),
\quad {\rm for}\ i\in\{0,1,2\}.
\label{eq:rate_density_shape}
\end{equation}
The shape function is normalized to unity.
The $Z_i$'s can be interpreted as each system type's occurrence rate 
$\Lambda_i$, integrated over all planetary radii.
They are equivalent to the number of planets per single star, primary, or 
secondary.

\subsection{Searchable stars}
For a given planet radius, semimajor axis, and stellar luminosity, a fixed 
signal-to-noise floor yields a maximum searchable 
distance~\citep{pepper_using_2003,pepper_searching_2005}.
From Eq.~\ref{eq:S_N_thresh}, this maximum searchable distance scales as
\begin{equation}
d(\delta_{\rm obs}, L_{\rm sys}) \propto \delta_{\rm obs} \cdot L_{\rm 
    sys}^{1/2}.
\end{equation}
Assuming that stars are uniformly distributed in space, the number 
of searchable stars $N\s$ is then proportional to
\begin{equation}
N\s(\delta_{\rm obs}, L_{\rm sys}) \propto n \delta_{\rm obs}^3 L_{\rm 
    sys}^{3/2},
\label{eq:N_searchable_prop}
\end{equation}
where $n$ is the volume density of {\it e.g.}, single star, or binary systems.
We neglect the dependence of $n$ on stellar type.

\subsection{Relation between Apparent and Actual Stellar Properties}

We assume that the apparent properties of an unresolved binary are identical to
the true properties of the primary:
\begin{equation}
M\a=M_1, \quad R\a=R_1, ...
\end{equation}
We also take the stellar radius and luminosity to be uniquely related to the 
stellar mass.
Given these assumptions, the total luminosity of a system is
\begin{equation}
L_{\rm sys}=L_1+L_2=L(M\a)+L(qM\a),
\end{equation}
where $q=M_2/M_1$. Note that $L_{\rm sys}$ is the true value (because $M\a$ is 
the true primary mass), while an observer with no a priori knowledge of the 
distance to a system would measure a flux, assume a stellar mass $M_a$, and 
estimate a system luminosity of $L(M\a)$.

\subsection{Apparent number of searchable stars}
\label{subsec:apparent_N_searchable}

\begin{figure}[!tb]
    \centering
    \includegraphics[width=0.6\textwidth]{figures/mass_ratio_distribution.pdf}
    \caption{
        The mass ratio distribution for a magnitude-limited sample of 
        binary stars, in which the underlying volume-limited distribution is 
        uniform, qualitatively similar to {\it e.g.}, 
        \citet{raghavan_survey_2010}'s Fig.~16.
        At a given observed depth, the searchable binaries in a transit survey 
        are magnitude-limited (see Sec.~\ref{subsec:apparent_N_searchable}).
    }
    \label{fig:q_distribn_mag_limited}
\end{figure}

Given the observed signal $\delta_{\rm obs}$ and apparent stellar mass $M\a$, 
the maximum searchable distance for singles and binaries are proportional to 
$\delta_{\rm obs}\cdot L(M\a)^{1/2}$ and $\delta_{\rm obs}\cdot 
[L(M\a)+L(qM\a)]^{1/2}$.
Thus, the apparent number of searchable stars (points in the sky), which 
will be selected by an ignorant observer, can be written as
\begin{align}
N_{\rm s,a}(\delta_{\rm obs}, M\a)
= N\s^0(\delta_{\rm obs}, L(M\a)) \left[1+\mu(\mathrm{BF}, 
M\a)\right],
\label{eq:N_selected}
\end{align}
where $N\s^0$ is the number of searchable singles (this agrees with the actual 
value), $\mathrm{BF}=n_{\rm b}/(n_{\rm s}+n_{\rm b})$ 
is the binary fraction in a volume-limited sample,
$n_s$ ($n_b$) is the number density of single (binary) systems in a 
volume-limited sample,
and $\mu$ is the ratio of the number of double to single star 
systems that are searchable for an observed signal $\delta_{\rm obs}$.
Applying Eq.~\ref{eq:N_searchable_prop},
\begin{equation}
\mu({\rm BF},M\a) = 
{N_{\rm d} \over N_s^0}=
\int_0^1 {n_{\rm b}\over n_{\rm s}}\left[1+{L(qM\a) \over L(M\a)}\right]^{3/2} 
f(q)\mathrm{d}q,
\label{eq:mu_general}
\end{equation}
where $f(q)$ is the distribution of binary mass ratios in a volume-limited 
sample.
If $\mathrm{BF}$ is independent of $q$, $L \propto M^\alpha$, and $f(q) 
\propto q^\beta$, this reduces to
\begin{equation}
\mu =\frac{{\rm BF}}{1 - {\rm BF}}\cdot {1\over{1+\beta}}\int_0^1 
\left(1+q^\alpha\right)^{3/2} q^\beta\mathrm{d}q,
\end{equation}
which may be written in a closed form using the hypergeometric function.

This can be understood as a Malmquist bias: at fixed observed transit 
depth, for singles and primaries with the same luminosity, binaries are 
searchable to a greater distance.
In particular, the sample of searchable binaries is magnitude-limited.
Given a searchable binary, the probability of drawing a mass ratio $q$ will 
scale as
\begin{equation}
p({\rm draw\ }q\,|\,{\rm system\ is\ binary}) \propto 
(1+q^\alpha)^{3/2} q^\beta 
\label{eq:Malmquist_bias}
\end{equation}
where $q^\beta$ is the volume-limited probability of drawing a binary of mass 
ratio $q$, and the first term is the Malmquist bias.
We show this magnitude-limited mass ratio distribution for the $\beta=0$ case 
in Fig.~\ref{fig:q_distribn_mag_limited}.
In Monte Carlo simulations of transit surveys, it is 
important to draw binaries from a correctly biased mass-ratio distribution 
\citep[\textit{e.g.},][]{bakos_hatsouth:_2013,sullivan_transiting_2015,
    gunther_new_2017}.

>>>>

\documentclass[12pt,modern]{aastex61}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{comment}
\usepackage{grffile} % checks for multiple dots in figure filenames
\usepackage{empheq} % for boxing equations
\newcommand*\widefbox[1]{\fbox{\hspace{0.2cm}#1\hspace{0.2cm}}}

@@

\newcommand{\pp}{\mathcal{P}}
\newcommand{\ps}{\mathcal{S}}
\renewcommand{\a}{_{\rm a}}
\newcommand{\s}{_{\rm s}}
\newcommand{\p}{_{\rm p}}
\renewcommand{\d}{_{\rm d}}
\begin{document}

\title{ The effects of binarity on planet occurrence rates measured by transit 
surveys}
%
\correspondingauthor{L. Bouma}
\email{luke@astro.princeton.edu}
%
\author{L. G. Bouma}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author{K. Masuda}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author{J. N. Winn}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
%
%
\begin{abstract}
%
This study develops simplified models of signal-to-noise limited transit 
surveys, in order to clarify the biases that stellar binarity introduces
to occurrence rate measurements.
Particular attention is paid to dilution of observed planet radii, errors in 
star counts, and detection efficiencies.
Free parameters in our models include the true radius distribution of planets, 
and the relative numbers of planets orbiting secondaries, primaries, and 
single stars.
In the simplest possible case, all stellar systems are either single or twin 
binaries, and all planets are identical. If so, then we find that ignoring 
binarity leads to an underestimate of the occurrence at the true radius by a 
multiplicative factor of $0.76$ (assuming a 10\% twin binary fraction).
Using more realistic models for the stellar population and planetary 
radii, we find that the occurrence of Earth-sized planets can be overestimated 
by at most $\approx 50\%$ when ignoring binarity~--~at present far smaller 
than systematic uncertainties on real $\eta_\oplus$ measurements.
We quantify how gaps in the intrinsic planet radius distribution are filled 
in when binarity is neglected.
Our models suggest that for most of the observed radius distribution 
($r>2r_\oplus$), ignoring binarity only leads to errors in inferred occurrence 
rates at the $\lesssim 10\%$ level.
One corollary is that binarity is unlikely to strongly influence the HJ rate 
discrepancy.
%
\end{abstract}
%
\keywords{
    methods: data analysis ---
    planets and satellites: detection ---
    surveys}
%
%

@@

To compute an occurrence, at every planetary and stellar property the 
astronomers choose the stars around which the planets of interest appeared to 
be searchable.
Correcting for the geometric transit probability $p_{\rm tra}$, they report an 
apparent rate density,
\begin{equation}
\Gamma\a(\pp\a, \ps\a) = \frac{n_{\rm det}(\pp\a, \ps\a)}{N_{\rm s,a}(\pp\a, 
    \ps\a)} \times \frac{1}{p_{\rm tra}(\pp\a, \ps\a)}.
\label{eq:defn_apparent_rate_density}
\end{equation}
where $\pp\a$, $\ps\a$ are the apparent planetary and stellar parameters.
The quantity $N_{\rm s,a}(\pp\a, \ps\a)$ is the number of unresolved 
point-sources on the sky that appear to be searchable, and
$n_{\rm det}(\pp\a, \ps\a)$ is the number of detected planets, per unit 
$\pp\a$ and $\ps\a$.

@@

An immediate complication is that, due to dynamical stability or some 
aspect of planet formation, the occurrence rate of planets might differ 
between binary and single-star systems.
In an observed sample of stars that contains both singles and binaries, the 
detected planets would then be drawn from different intrinsic 
distributions for the occurrence rates in single star systems, 
about primaries, and about secondaries~\citep[{\it 
e.g},][]{wang_occurrence_2015}.

Outside of astrophysical differences, every term in 
Eq.~\ref{eq:defn_apparent_rate_density} is observationally-biased.
There are errors in $n_{\rm det}(\pp\a,\ps\a)$ due to misclassification of 
both planet radii and stellar properties.
There are errors in $N_{{\rm s,a}}(\pp\a,\ps\a)$ because an apparently 
searchable point on the sky might include two searchable stars.
There are errors in $p_{\rm tra}(\pp\a,\ps\a)$ because stars in binaries may 
have different masses and radii than assumed in the single-star case.
Binarity might also affect the observers' ability to correctly select 
searchable stars (but see Sec.~\ref{sec:discussion} for discussion of this 
issue).

Correcting for binarity's observational biases is non-trivial.
For instance, in counting the number of searchable stars, even after realizing 
that binaries count as two stars, one must note that the multiplicity fraction 
of stars that are searchable for a given observed transit depth is greater 
than that the multiplicity fraction in a volume limited sample.
This is the familiar Malmquist bias: at a fixed $\delta_{\rm obs}$, binaries 
are searchable out to larger distances than single stars because they are more 
luminous.
While this is what biases detectability, separate effects bias interpretation:
ignoring binarity, the (apparent) radii assigned to planets in binaries will 
typically be smaller than the true radii.
The apparent and true radii differ because of diluting flux, and possibly 
because the planet is assumed to orbit the wrong 
star.

To gain intuition for the many observational biases at play,
we consider a set of idealized transit surveys.
First, we discuss the simplest possible cases (Sec.~\ref{sec:simplest}), in 
which all planets have identical properties, and all stars have identical 
properties except that some are in binaries.
The straightforward thought-experiments of Sec.~\ref{sec:simplest} are then 
generalized in Sec.~\ref{sec:general_formula} to allow for varying populations 
of secondaries, and for varying intrinsic occurrence rate distributions.
Sec.~\ref{sec:more_complicated} investigates questions of interest, 
including (a) measurements of $\eta_\oplus$, (b) the ``hot Jupiter rate 
discrepancy'', and (c) \citet{fulton_california-_2017}'s 
recently discovered ``valley''.
We discuss what our models suggest about binarity's biasing effects in 
Sec.~\ref{sec:discussion}, and conclude in Sec.~\ref{sec:conclusion}.

@@

\section{The simplest possible transit survey models}
\label{sec:simplest}

\subsection{Twin binaries, identical stars, identical planets}
\label{sec:model_1}

\begin{figure}[!tb]
    \begin{center}
        \includegraphics[width=0.6\textwidth]{figures/visualize_volumes.pdf}
    \end{center}
    \caption{
        Cartoon of the searchable volumes for single stars ({\it left}), and 
        twin binaries ({\it right}).
        This model assumes all stars have equal mass and luminosity, and 
        all planets have the same true radius $r=r\p$.
        {\it Top:} At an apparent radius $r\a=r\p$,
        the observer searched twin binaries out to a larger         
        distance, $\sqrt{2}\times$ that of single stars.
        Because of dilution, there are no planets in twin binaries with 
        $r\a=r\p$.
        {\it Middle:} At an apparent radius $r\a=r\p/\sqrt{2}$, the 
        searched volumes are half those at $r\a=r\p$. 
        The only detected planets with $r\a=r\p/\sqrt{2}$ orbit twin binaries.
        {\it Bottom:} Planets with a true radius $r=r\p$ are searchable to 
        a maximum distance $d_0$ around singles, and $d_0/\sqrt{2}$ around 
        twin binaries.
    }
    \label{fig:model_1_volumes}
\end{figure}

Since the effects of binarity are most pronounced when the two components are 
similar, we begin by considering a universe in which all planets are 
identical, and all stars are identical except that some fraction of them exist 
in binaries. The goal of this thought experiment is to compare the occurrence 
inferred by observers ignoring binarity with the true occurrence in single 
star systems.

First, what are the possible apparent radii of detected planets?
If the true planet radius is $r=r\p$, then planets in singles will be 
detected with the identical apparent radii: $r\a = r\p$.
Conversely, all detections in twin binaries will have apparent radii 
$r\a=r\p/\sqrt{2}$, because they are diluted.
For an unresolved point-source,
\begin{equation}
\delta_{\rm obs}
= \left(\frac{r\a}{R\a}\right)^2
= \left[{r\over R_{\rm host}}\right]^2\times {L_{\rm host} \over L_{\rm 
        sys}(M_1, q)},
\label{eq:delta_obs_general} 
\end{equation}
where $R_{\rm host}$ ($L_{\rm host}$) is the planet-host's stellar radius 
(luminosity), $R\a$ is the apparent stellar radius, $L_{\rm sys}=L_1+L_2 = 
L(M_1) + L(qM_1)$ is the system luminosity, and $q=M_2/M_1$ is the binary mass 
ratio.
We assume that the stellar radius and luminosity are uniquely related to the 
stellar mass.
The observers, with no a priori knowledge of the distance to a system, would 
measure a flux, assume a stellar mass $M\a=M_1$, and estimate both a stellar 
radius $R\a = R(M\a)$ and an under-luminous system luminosity of $L_{\rm 
sys,a}(M\a) = L(M\a)$.

The planets will be detected if and only if their host stars are searchable.
For a fixed signal-to-noise floor, there is a maximum searchable 
distance~\citep{pepper_using_2003,pepper_searching_2005}.
From Eq.~\ref{eq:S_N_thresh}, this maximum searchable distance scales as
\begin{equation}
d(\delta_{\rm obs}, L_{\rm sys}) \propto \delta_{\rm obs} \cdot L_{\rm 
    sys}^{1/2}.
\end{equation}
Assuming that stars are uniformly distributed in space, the number 
of searchable stars $N\s$ is then proportional to
\begin{equation}
N\s(\delta_{\rm obs}, L_{\rm sys}) \propto n \delta_{\rm obs}^3 L_{\rm 
    sys}^{3/2},
\label{eq:N_searchable_prop}
\end{equation}
where $n$ is the number per unit volume of {\it e.g.}, single star, or binary 
systems.
We neglect the dependence of $n$ on stellar type.

The searchable volumes for our thought experiment are illustrated 
in Fig.~\ref{fig:model_1_volumes}.
At an apparent radius equal to the true radius ($r\a=r\p$), single stars 
with $d<d_0$ are searchable.
Binaries with $r\a=r\p$ are also searchable, out to $\sqrt{2}d_0$. 
However, in this model no such systems exist; all planets in twin binaries 
have true radii $r=r\p$, and so are observed with apparent radii 
$r\a=r\p/\sqrt{2}$, out to $d_0/\sqrt{2}$.
At an apparent radius of $r\a=r\p/\sqrt{2}$, there could also be planets with 
true radii $r=r\p/\sqrt{2}$ orbiting singles~---~however none of these exist 
in this model.

How many planets are detected?
If we assume that there are $N\s^0(r\a)$ singles that are searchable for 
planets with $r\a$, and that there are $Z_0$ planets per single, then 
single stars contribute
\begin{equation}
n_{\rm det}^0(r\a) = N\s^0(r\a) Z_0 p_{\rm tra} \cdot \delta(r\a - r_p)
\end{equation}
planet detections (per unit $r\a$), where $\delta$ is the Dirac delta function.

We can write similar equations for the binaries.
First though, we define a ratio $\mu \equiv N\s^{\rm b}(r\a) / N\s^0(r\a)$ 
between the number of searchable binary systems and the number of searchable 
singles at a given $r\a$.
This quantity is useful because although $N\s^{\rm b}$ and $N\s^0$ vary
as a function of the observed transit depth, $\mu$ remains fixed.
Applying this definition, at $r\a=r\p/\sqrt{2}$, some detections will come 
from primaries,
\begin{equation}
n_{\rm det}^1(r\a) =
\mu N\s^0(r\a) Z_1 p_{\rm tra}
\cdot \delta\left(r\a - \frac{r\p}{\sqrt{2}}\right),
\end{equation}
and some from secondaries,
\begin{equation}
n_{\rm det}^2(r\a) = 
\mu N\s^0(r\a) Z_2 p_{\rm tra}
\cdot \delta\left(r\a - \frac{r\p}{\sqrt{2}}\right),
\end{equation}
where $Z_1$ is the number of planets per primary, $Z_2$ is the number of 
planets per secondary, and $p_{\rm tra}$ is the same for identical stars.
In the case of twin binaries, Eq.~\ref{eq:N_searchable_prop} gives $\mu = 
2^{3/2} n_{\rm b}/n_{\rm s} = 2^{3/2} {\rm BF}/(1-{\rm BF})$, for ${\rm BF}$ 
the binary fraction\footnote{The binary fraction is the fraction of systems in 
a volume-limited sample that are binary. It is equivalent to the multiplicity 
fraction if there are no triple, quadruple, or higher order multiples.
}.

The last item we need to write down the apparent rate density $\Gamma\a(r\a)$ 
is the number of unresolved point-sources on sky, $N_{\rm s,a}$.
Assuming that the apparent stellar radius is fixed, then given the 
apparent planet radius $r\a$,
\begin{equation}
N_{\rm s,a}(r\a) = N\s^0(r\a) \times (1 + \mu)
\label{eq:apparentlysearchable_to_searchable}
\end{equation}
relates the number of apparently searchable point-sources to the number of 
searchable singles.
From the discussion above, it should be clear that there are in fact
$N\s^0(r\a)\times (1 + 2\mu)$ searchable stars at any given $r\a$; the 
observers are under-counting.

Using Eq.~\ref{eq:defn_apparent_rate_density}, the apparent rate density 
computed without considering binarity is
\begin{align}
\Gamma\a(r\a) = 
\frac{1}{1+\mu} Z_0 \cdot
\delta(r\a-r\p)  +
\frac{\mu}{1+\mu} (Z_1 + Z_2) \cdot
\delta\left(r\a-\frac{r\p}{\sqrt{2}} \right).
\label{eq:model_1_apparent_rate_density}
\end{align}
The number of searchable singles at a given apparent radius, 
$N\s^0(r\a)$, makes no appearance.
The observer's errors in Eq.~\ref{eq:model_1_apparent_rate_density} are 
(A) under-counting the number of searchable stars at any apparent radius;
and
(B) inferring radii in binary systems that are all $\sqrt{2}$ too small.

\begin{figure}[!tb]
    \begin{center}
        \includegraphics[width=.6\textwidth]{figures/occ_rate_vs_radius_model_1_brokenx.pdf}
    \end{center}
    \vspace{-0.5cm}
    \caption{
        Inferred occurrence rates over $0.01r_\oplus$ bins in planet 
        radius, for a model with identical stars, identical planets, and twin 
        binaries.
        We assume a twin binary fraction of ${\rm BF}=0.1$, and that all 
        stars (single, primary, and secondary) host planets at equal 
        rates.
        If the true planet radius is $r\p$, all planets detected in binaries 
        will have apparent radii $r\a = r\p/\sqrt{2}$.
    }
    \label{fig:occ_rate_model_1}
\end{figure}

To assess the severity of these errors, we need to assume a binary fraction, 
and to know the true number of planets per single, primary, and secondary star 
($Z_0,Z_1,$ and $Z_2$).
For the former,
\citet{raghavan_survey_2010} found a multiplicity 
fraction of 0.44 for primaries with masses from $0.7M_\odot$ to $1.3M_\odot$, 
which we take as a binary fraction. 
``Twin binaries'' are perhaps 10-20\% of the binaries in a volume-limited 
sample, depending on how ``twin'' is defined.
Thus a representative twin binary fraction for this model is ${\rm BF}\approx 
0.05$-$0.1$. 
For the latter, we take $Z_0=Z_1=Z_2=0.5$, and plot the resulting occurrence 
rates as a function of planet radius in Fig.~\ref{fig:occ_rate_model_1}.

Our point for comparison is the rate density for single stars,
\begin{equation}
\Gamma_0(r) = Z_0 \cdot \delta(r-r\p).
\end{equation}
We can define a rate density ``correction factor'', $X_\Gamma$, as the ratio 
of the apparent to single star rate density,
\begin{equation}
X_\Gamma(r) \equiv \left. \frac{\Gamma_a(r\a)}{\Gamma_0(r)} \right|_{r\a 
    \rightarrow r}.
\end{equation}
Continuing in the assumption that the number of planets per single, primary, 
and secondary star are equal, we find a rate density 
correction factor at the true planet radius of $1/(1+\mu)$.
This yields a correction of 0.76 if ${\rm BF}=0.1$, and 0.87 if ${\rm 
    BF}=0.05$. Rephrasing the result, if the twin binary fraction were 
$(2^{3/2}+1)^{-1}\approx 0.26$, then the apparent rate at $r\a=r\p$ would be 
half the true rate at $r=r\p$.
For most stellar types, the twin binary fraction is not that high.

An alternative assumption is that secondaries do not host planets. In that 
case, $Z_0=Z_1$, and $Z_2=0$. The correction to the rate density at the true 
planet radius becomes $(1+2\mu)/(1+\mu)^2$.
This evaluates to 0.94 if ${\rm BF}=0.1$, and 0.98 if ${\rm BF}=0.05$.
While it seems unlikely that secondaries are planet-less, 
it is worth noting that $\Gamma\a/\Gamma$ is sensitive to the relative number 
of planets per single, primary, and secondary.

\subsection{Twin binaries, identical stars, two planet sizes}

A simple extension to the previous example will help us further distinguish 
the detected populations.
Consider now a universe that is the same in every respect to that of 
Sec.~\ref{sec:model_1}, except now half of planets have radii $r\p$, while 
the other half have $r\p/\sqrt{2}$.
The true rate densities are
\begin{equation}
\Gamma_i(r) = \frac{Z_i}{2} \left[
\delta(r-r\p) + \delta\left(r-\frac{r\p}{\sqrt{2}}\right)
\right], \quad {\rm for}\  i \in \{ 0,1,2 \},
\end{equation}
where as always $i=0$ corresponds to singles, $i=1$ to primaries, and $i=2$ to 
secondaries, and the $Z_i$'s are the number of planets per star of each type.

Following an identical line of reasoning as before, the apparent rate density 
can be written
\begin{align}
\notag
\Gamma\a(r\a) &=
\frac{1}{2(1+\mu)} \left[
Z_0 \cdot \delta(r\a - r\p)
+
\left[Z_0 + \mu (Z_1 + Z_2)
\right] \cdot \delta\left(r\a - \frac{r\p}{\sqrt{2}}\right) 
\right. \\
&\quad\quad\quad\quad\quad\quad\quad\quad+
\left.
\mu(Z_1 + Z_2) \cdot \delta\left(r\a - \frac{r\p}{2}\right)
\right].
\label{eq:two_radii_twins}
\end{align}
The important qualitative point of this equation is that at $r\a = 
r\p/\sqrt{2}$, two populations contribute.
The first population is singles with $r=r\p/\sqrt{2}$, and the second is twin 
binaries with $r=r\p$.
Both populations are detected out to the same maximum searchable distance.
Just as with Eq.~\ref{eq:apparentlysearchable_to_searchable}, we will have a 
cancellation of $N\s^0(r\a)$ terms, leaving only the $\mu$ weights.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{General formula for apparent occurrence rate}
\label{sec:general_formula}

To generalize the procedure of Sec.~\ref{sec:simplest}, we consider a S/N 
limited transit survey in which all the primaries and single stars have 
identical properties.
For example, they might all be solar analogs, roughly similar to 
the stars that {\it Kepler} surveyed~\citep{batalha_selection_2010}.
We let the properties of the secondaries vary, and assume that there are some 
functions $L(M)$ and $R(M)$ that specify a star's luminosity and radius in 
terms of its mass.
We then allow for an arbitrary volume-limited mass ratio distribution 
$f(q)$, and also for arbitrary true rate densities for singles, primaries, 
and doubles ($\Gamma_0, \Gamma_1, \Gamma_2$).
Since the properties of the secondaries vary, we must consider not only the 
dependence of the rate densities on planet radius, but also on stellar mass.
We presume that the observers always take the properties of a binary system to 
be those of the primary.

Given this laundry-list of assumptions, a general formula for the apparent 
rate density follows:
\begin{align}
\notag
\Gamma\a(r\a,M\a) &= {1\over 1+\mu(\mathrm{BF}, M\a)}\times
\left\{ \Gamma_0(r\a, M\a)+ 
\frac{{\rm BF}}{1 - {\rm BF}}
\left[ \int {\mathrm{d}q \over \mathcal{A}^4}\,f(q)\,\Gamma_1\left({r\a\over 
    \mathcal{A}}, 
M\a\right)\,
\right.   
\right. \\
& \quad\quad\quad\quad\quad \left.\left.
+\int {q \mathrm{d}q \over 
    {\mathcal{A}^3\mathcal{B}}}\,f(q)\,\Gamma_2\left({r\a\over 
    \mathcal{B}}, 
qM\a\right)\,
{R(qM\a) \over R(M\a)}
q^{-1/3} \right]	\right\},
\label{eq:general_Gamma_a}
\end{align}
where 
\begin{equation}
\mathcal{A}(q, M\a)=\sqrt{L(M\a) \over L_{\rm sys}(M\a, q)}
\end{equation}
and
\begin{equation}
\mathcal{B}(q, M\a)={R(M\a)\over R(qM\a)}\sqrt{{L(qM\a) \over L_{\rm 
            sys}(M\a, q)} }
\end{equation}
are the dilution terms appropriate to primaries and secondaries, and as before 
${\rm BF}$ is the volume-limited binary fraction.
